### 2023-01-31 scripts/01 Filter Bioeconomy.qmd
add scripts to upload tags to db

---
title: "01 Filter Forest Bioeconomy"
format: html
---
```{r}
here::i_am("scripts/01 Filter Bioeconomy.qmd")

library(tidyverse)
library(here)
library(DataEditR)

data_path <- here("data")

```

```{r}
#| warning: false

df <- readxl::read_xlsx(here(data_path, "raw-data", "20201221 SWINNO 1970-2019.xlsx"), sheet="SWINNO 1970-2019")

additional_info <- df %>% select(starts_with("Additional")) %>% colnames()
use_sectors <- df %>% select(starts_with("Use")) %>% colnames()

df <- df %>% 
  unite(info, additional_info, sep =" ", remove=TRUE, na.rm=TRUE) %>%
  unite(sectors, use_sectors, sep=",", remove= TRUE, na.rm=TRUE)

df <- df %>%
  select(c(`SINNO ID`, `Literal name of innovating firm ; SINNO's common name ; SINNO ID`  ,`Year of commercialization`, `Description in Swedish`, info, sectors))

filtered_df  <- df %>%
  filter(sectors == 2) %>%
  rename(year = `Year of commercialization`, description = `Description in Swedish`)

write_excel_csv(filtered_df, here(data_path, "modified-data", "classification-forest.csv"))

```

```{r}
data\raw-data\

df <- read_csv(here(data_path, "raw-data", "chaminade-bayou-bioeconomy.csv"))

additional_info <- df %>% select(starts_with("Additional")) %>% colnames()

df <- df %>% 
  unite(info, additional_info, sep =" ", remove=TRUE, na.rm=TRUE)

df <- df %>%
  select(c( SINNO.ID,
        Name.Swe,
        Descri.Swe,
        Literal.name.of.innovating.firm...SINNO.s.common.name...SINNO.ID,
       info))

write_excel_csv(df, here(data_path, "modified-data", "classification-bioeconomy.csv"))

```

```{python}

import pandas as pd

df = pd.read_csv(
    "../data/raw-data/chaminade-bayou-bioeconomy.csv",
    usecols=[
        "SINNO.ID",
        "Name.Swe",
        "Descri.Swe",
        "Literal.name.of.innovating.firm...SINNO.s.common.name...SINNO.ID",
        "Additional.information.if.Origin...1a.New.scientific.discovery",
        "Additional.information.if.Origin...1b.New.technologies.or.materials",
        "Additional.info.if.Origin...2a.Official.regulation..legislation.and.standards.",
        "Additional.information.if.Origin...5b.Solution.for.a.problem",
        "Additional.information.if.Origin...7.Performance",
        "Additional.information.if.Origin...20.Other",
    ],
)

add_info = list(df.iloc[:, 4:].columns)

df["info"] = df[add_info].apply(lambda x: x.str.cat(sep=''), axis=1)

df.to_csv("../data/modified-data/bioeonomy-classification.csv")

```