### 2023-09-25 docs/presentations/20230928-sehm_presentation.qmd

---
title: Shades of Green
subtitle: Direction of the Swedish Forest Innovation System 1970-2020
author: Jonas Kreutzer
date: last-modified
abstract: |

csl: https://zotero.org/styles/apa?source=1
bibliography: [../../swinno-shades-green.bib]

format:
  revealjs:
    theme: [../assets/revealme.scss]
    transition: none
    title-slide-attributes:
      data-background-image: '../assets/ground-fog-valley-black-forest-sunset.jpg'
      data-background-size: fit
      data-background-opacity: '0.3'
    embed-resources: true

execute:
    freeze: auto
    echo: false
jupyter: python3
---

```{python}
# | label: Imports

import pandas as pd
import numpy as np
from IPython.display import display, Markdown
import matplotlib.pyplot as plt
import matplotlib.ticker as mtick
import seaborn as sns

from src.figures import *
from src.utils import get_project_root, format_table
from src.swinno_helpers import connect_swinno_db

ROOT = get_project_root()
conn = connect_swinno_db()

# set all text on plots to be monospace with rcParams
plt.rcParams["font.family"] = "Space Grotesk"

sns.set_theme(style="white")

# full code labels for merging

codes_df = pd.read_sql("SELECT * FROM classification_codes", conn)

codes_df["Category"] = codes_df["Category"].str.strip()
uncertain_eco = {"Code": "000", "Category": "Unsure"}
codes_df = codes_df.append(uncertain_eco, ignore_index=True)

codes_df = codes_df[~codes_df["Code"].duplicated(keep="last")]

codes_df = codes_df.set_index("Code")
```

```{python}
# | label: Query bioeconomy

bioeconomy = pd.read_sql(
    """
select
  i.sinno_id,
  i.innovation_name_in_swedish AS name,
  i.description_in_swedish AS description,
  i.additional_information_if_origin_new_scientific_discovery || i.additional_information_if_origin_new_technologies_or_materials || i.additional_info_if_origin_official_regulation_legislation_and_standards || i.additional_information_if_origin_solution_for_a_problem || i.additional_information_if_origin_performance || i.additional_information_if_origin_other AS info,
  i.year_of_commercialization AS year,
  us.use_sector
from
  innovation i
  join use_sectors us on i.sinno_id = us.sinno_id
where
  (
    us.use_sector like '02%'
    or us.use_sector like '20%'
    or us.use_sector like '21%'
    or us.use_sector like '36%'
    or product_code like '02%'
    or product_code like '20%'
    or product_code like '21%'
    or product_code like '36%'
  )
  or (
    description like '%virke%'
    or description like '%cellulos%'
    or description like '%lignin%'
    or description like '%spån%'
    or description like '%bark%'
    or description like '%levulinsyra%'
    or description like '%furfural%'
    or description like '%svarttjära%'
    or description like '%svartlut%'
    or description like '%växtbas%'
    or description like '%ved%'
    or description like '%trä%'
    or description like '%skog%'
    or description like '%biobränsle%'
    or description like '%biologisk%'
    or description like '%nedbrytbar%'
    or description like '%papper%'
    or description like '%pappret%'
    or description like '%karton%'
    or description like '%tencel%'
  );
""",
    conn,
)

sni_codes = pd.read_sql_query(
    """
select
*
from 
sni_codes
""",
    conn,
)

sni_codes = sni_codes.rename(columns={"code": "use_sector"})

swinno = pd.read_sql_query(
    """
select sinno_id, year_of_commercialization as year, innovation_name_in_swedish as name
from innovation;
""",
    conn,
)

swinno["bioeconomy"] = swinno["sinno_id"].isin(bioeconomy["sinno_id"].unique())

```

```{python}
# | label: Query eco-innovation

eco_innovations = pd.read_sql(
    """
SELECT
  sinno_id,
  innovation_type as innovation_type_code,
  c.Category as innovation_type
FROM
  eco_innovations
JOIN classification_codes AS c ON c.Code = eco_innovations.innovation_type
WHERE 
innovation_type NOT IN (701, 602, 601, 501, 111, 109, 108)
;
""",
    conn,
)

eco_innovations["innovation_type"] = eco_innovations["innovation_type"].str.strip()

```

```{python}
# | label: Calculate Bioeconomy Share

def calculate_bioeconomy_share(swinno):
    df = swinno.copy()
    df = df[["sinno_id", "year", "bioeconomy"]]
    df = (
        df.groupby("year")
        .agg(
            total_count=("sinno_id", "nunique"), bioeconomy_count=("bioeconomy", "sum")
        )
        .reset_index()
    )
    df["bioeconomy_share"] = df["bioeconomy_count"] / df["total_count"]

    return df

bioeconomy_share = calculate_bioeconomy_share(swinno)
```

## Bioeconomy Demarkations {.smaller}

::::{.columns}

:::{.column width="50%"}
**Using producer / user codes
**

|SNI Code | Description                            |
|--------:|:---------------------------------------|
|02       | Forestry and related services
|20       | Wood and wood product manufacturing except furniture
|21       | Pulp, paper and paper product manufacturing
|36       | Furniture manufacturing; other manufacturing
:::

:::{.column width="50%"}
**Adding keywords of value chain**

    - "virke",
    - "cellulos",
    - "lignin",
    - "spån", 
    - "bark",
    - "levulinsyra" (Levulinic acid),
    - "furfural" (Furfural),
    - "svarttjära",
    - "svartlut",
    - "växtbas",
    - "ved",
    - "trä",
    - "skog",
    - "papper",
    - "biobränsle",
    - "biologiskt",
    - "nedbrytbar",
    - "papper",
    - "pappret",
    - "karton",
    - "tencel",
:::

::::

****

## 817 Innovations in Forest Bioeconomy (Uncleaned)

## Count of Bioeconomy Innovation
### [Yearly]{.text-royal-blue-800} and [5 Year MA]{.text-gray-600}
```{python}
# label: fig-bioeconomy_count
#
#

plot_bioeconomy_trends_interactive(
    bioeconomy_share, y="bioeconomy_count"
).update_layout(yaxis_range=[0, 35]).show()

```

## Share of Bioeconomy Innovation{.smaller}
### [Yearly]{.text-royal-blue-800} and [5 Year MA]{.text-gray-600} 
```{python}

plot_bioeconomy_trends_interactive(bioeconomy_share, y="bioeconomy_share").update_yaxes(
    range=[0, 0.35], tickformat=".0%"
).show()

```

# Three Visions for a Bioeconomy

:::: columns
::: {.column width="33%"}
![Bio-Technology Vision](../assets/Biotech.jpg){.fragment}
:::

::: {.column width="33%"}
![Bio-Resource Vision](../assets/Bioresource.jpg){.fragment}
:::

::: {.column width="33%"}
![Bio-Ecology Vision](../assets/Bioecology.jpg){.fragment}
:::
::::

[@bugge2016WhatBioeconomyReview]{.slide-cite style="font-size: x-small"}

## Bioeconomy Vision Classification{.smaller .scrollable}

| @bugge2016WhatBioeconomyReview | Biotechnology | Bioresource | Bioecology |
|---|---|---|---|
| Aim &   Objectives | Economic growth & job   creation | Economic growth &   sustainability | Sustainability, biodiversity,   conservation of ecosystems, avoiding soil degradation |
| Value   Creation | Application of biotechnology,   commercialisation of research & technology | Conversion and upgrading of   bio-resources (process oriented) | Development of integrated   production systems and high-quality products with territorial identity |
| Drivers   & mediators of innovation | R & D, patents, TTOs,   Research councils and funders (Science push, linear model) | Interdisciplinary, optimisation   of land use, include degraded land in the production of biofuels, use and   availability of bio-resources, waste management, engineering, science &   market (Interactive & networked production mode) | Identification of favourable   organic agro-ecological practices, ethics, risk, transdisciplinary   sustainability, ecological interactions, re-use & recycling of waste,   land use, (Circular and self-sustained production mode) |

| @vivien2019HijackingBioeconomy | Type II | Type III | Type I |
|---|---|---|---|
|  | A science-based economy driven   by industrial biotechnology<br>     <br>     The cell is a factory | Biomass replaces fossil fuels   and mining to produce energy and materials<br>     <br>     Biorefining at the heart of ecological transition (multilevel perspective). | An ecological economy, that is   compatible with the biosphere <br>     <br>     Counter-expertise rather than concrete technical solutions Criticism from   social groups who remain at the margins of decisionmaking centers |
| Codes | 1 | 2 | 3 |

## Preliminary Results

```{python}
bv_df = pd.read_sql(
    """SELECT bv.sinno_id, bv.bioeconomy_vision, i.year_of_commercialization AS "i.year"
FROM bioeconomy_visions AS bv
JOIN innovation AS i ON i.sinno_id = bv.sinno_id
WHERE bv.sinno_id NOT IN (SELECT sinno_id FROM categorization_notes WHERE notes NOT LIKE "%not forest%")
OR bv.sinno_id NOT IN (SELECT sinno_id FROM bioeconomy_visions WHERE bioeconomy_vision <> 0);
""",
    conn,
)
```

::::{.columns}
:::{.column width='80%'}
```{python}

display(
    Markdown(
        f"<span class='text-primary' style='font-size:15rem; font-weight:900'>{bv_df.loc[:, 'sinno_id'].nunique()}</span>"
    )
)

```

:::

:::{.column width='20%'}
unique forest bioeconomy innovations classified
:::

::::

## Examples

## Bioeconomy Vision by Year

```{python}

bv_year = pd.read_sql(
    """
SELECT
  codes.Category AS Vision,
  COUNT(bioeconomy_vision) AS Count,
  i.year_of_commercialization AS Year
FROM
  bioeconomy_visions
  JOIN classification_codes AS codes ON codes.code = bioeconomy_visions.bioeconomy_vision
  JOIN innovation AS i on i.sinno_id = bioeconomy_visions.sinno_id
WHERE
  NOT EXISTS (
    SELECT
      *
    FROM
      bioeconomy_visions AS bv2
    WHERE
      bv2.sinno_id = bioeconomy_visions.sinno_id
      AND bv2.bioeconomy_vision = 0
  )
  AND bioeconomy_visions.sinno_id NOT IN (
    SELECT
      sinno_id
    FROM
      categorization_notes
    WHERE
      notes NOT LIKE "%not forest%"
  )
GROUP BY
Year, Vision
ORDER BY
  Year ASC,
  Count DESC;
  """,
    conn,
)

def plot_bv_trends(df, title):
    sns.set_style("white")
    chart = sns.lineplot(
        data=df,
        x="Year",
        y="Count",
        hue="Vision",
        palette="colorblind",
        linewidth=2.5,
    )
    chart.set_title(title, fontdict={"fontsize": 16, "fontweight": "bold"})
    chart.set_xlabel("Year", fontdict={"fontsize": 12, "fontweight": "bold"})
    chart.set_ylabel("Count", fontdict={"fontsize": 12, "fontweight": "bold"})
    chart.tick_params(labelsize=10)
    plt.legend(title="Vision", fontsize=10, title_fontsize=12)
    plt.tight_layout()

    # Customize x-axis tick labels
    years = df["Year"].unique()
    chart.set_xticks(years[::5])
    chart.set_xticklabels(years[::5], rotation=45, ha="right")

    plt.show(chart)

plot_bv_trends(bv_year, title="")

```

## Bioeconomy Vision Count by Sector Excluding Uncertains {.smaller}
```{python}
bv_sectors = pd.read_sql(
    """
SELECT
  c.Category AS vision,
  sni_codes.label AS sector,
  count(bv.bioeconomy_vision) AS count
FROM
  bioeconomy_visions AS bv
  join use_sectors ON bv.sinno_id = use_sectors.[sinno_id]
  join sni_codes ON SUBSTR(use_sectors.use_sector, 1, 2) = sni_codes.code
  join [classification_codes] AS c ON bv.bioeconomy_vision = c.Code
WHERE
  NOT EXISTS(
    SELECT
      *
    FROM
      bioeconomy_visions AS bv2
    WHERE
      bv2.sinno_id = bv.sinno_id
      AND bv2.bioeconomy_vision = 0
  )
  AND bv.sinno_id NOT IN (
    SELECT
      sinno_id
    FROM
      categorization_notes
    WHERE
      notes NOT LIKE "%not forest%"
  )
GROUP BY
  sector,
  bioeconomy_vision;
""",
    conn,
)

bv_matrix = prepare_heatmap_data(
    bv_sectors, index_col="sector", cols_col="vision", values_col="count"
)

plot_heatmap(bv_matrix)
```

## Uncertain Cases

```{python}
uncertains = pd.read_sql(
    """
SELECT
DISTINCT
  i.sinno_id,
  i.innovation_name_in_swedish AS name,
  i.description_in_swedish AS description,
  i.additional_information_if_origin_new_scientific_discovery || i.additional_information_if_origin_new_technologies_or_materials || i.additional_info_if_origin_official_regulation_legislation_and_standards || i.additional_information_if_origin_solution_for_a_problem || i.additional_information_if_origin_performance || i.additional_information_if_origin_other AS info,
  i.year_of_commercialization AS year
FROM
  innovation i
  JOIN use_sectors us ON i.sinno_id = us.sinno_id
  JOIN bioeconomy_visions bv on i.sinno_id = bv.sinno_id
WHERE i.sinno_id in (select distinct sinno_id from bioeconomy_visions where bioeconomy_vision = "0");
""",
    conn,
)

display(
    Markdown(
        f"{uncertains['sinno_id'].nunique()} cases are marked as unsure bioeconomy visions."
    )
)
```

## References