### 2024-09-05 20240906-pjk-NEST.qmd
add NEST presentation

---
title: Quantifying Directionality and Innovation Output in Sweden's Transition to a Forest-Based Bioeconomy
author: Philipp Jonas Kreutzer
date: last-modified
bibliography: swinno-shades-green.bib
csl: 'https://zotero.org/styles/apa?source=1'
format: 
    pjk_theme-revealjs
---

## [Forest-Based Bioeconomy]{.text-white} {background-image="/assets/images/forwarder.jpeg" .center}

## RQ {.center}

[1.]{.text-gray-600} [How many innovations are commercialized in this new economy]{.text-royal-blue-600}, especially considering the central role of innovation in bioeconomy discourse?

[2.]{.text-gray-600} What does commercialized innovation suggest for the [positive directionality of the bioeconomy transition]{.text-royal-blue-600}?

# Data From SWINNO Database {.inverted}

## Significant Swedish Innovations

::: {.columns}
::: {.column}
~ 5000 unique innovations

Literature-Based Innovation Output method

from 15 independent trade journals [@sjoo2014DatabaseSwedishInnovations]

:::
::: {.column}
![Example Page From a Source Article](/assets/images/swinno-example-1.jpg)
:::
:::

## Defining the Bioeconomy {.smaller}

::: {.columns}
::: {.column}
|SNI Code | Description                            |
|--------:|:---------------------------------------|
|02       | Forestry and related services
|20       | Wood and wood product manufacturing except furniture
|21       | Pulp, paper and paper product manufacturing
|36       | Furniture manufacturing; other manufacturing
:Key Sectors Used in Query {#tbl-query}

:::
::: {.column}
| Swedish           | English                    |
|-------------------|----------------------------|
| virke             | timber                     |
| cellulos          | cellulose                  |
| lignin            | lignin                     |
| spån              | chip                       |
| bark              | bark                       |
| levulinsyra       | levulinic acid             |
| furfural          | furfural                   |
| svarttjära        | black tar                  |
| svartlut          | black liquor               |
| växtbas           | plant-based                |
| ved               | wood                       |
| trä               | timber                     |
| skog              | forest                     |
| biobränsle        | biofuel                    |
| biologiskt        | biological                 |
| nedbrytbar        | biodegradable              |
| papper            | paper                      |
| karton            | carton                     |
| lyocell            | lyocell |
: Swedish Keywords used in Query: `WHERE description LIKE %keyword% OR ...` {#tbl-keywords}
:::
:::

# Results {.inverted}

## Bioeconomy Innovation Has Declined After 1970s

![a. A total number of innovation registered in SWINNO database. b. Forest-based bioeconomy innovations registered in SWINNO database. c. Percentage of forest based bioeconomy innovations to total innovation registered in SWINNO database. Shown are the 5 year moving averages for each time series.](results/figures/innovation_trends.svg){#fig-innovation-trends}

## Three Visions for Normative Directionality

::: {.columns}
::: {.column width="33%"}

### [Bioresource]{.text-batlow-darkgreen}

![](/assets/images/bioresource.svg)
:::
::: {.column width="33%"}

### [Biotechnology]{.text-batlow-forestgreen}

![](/assets/images/biotechnology.svg)

:::
::: {.column width="33%"}

### [Bioecology]{.text-batlow-darkblue}

![](/assets/images/bioecology.svg)

:::
:::

Classification based on @bugge2016WhatBioeconomyReview and @vivien2019HijackingBioeconomy

## Bioresource Was Biggest Vision Category

But most innovations did not align with any vision

![Counts of innovation classification by Bioeconomy Vision Category](/results/figures/vision_count.svg){#fig-bioeconomy-trends}

## [Bioresource Vision]{.text-batlow-darkgreen} {background-video="/assets/images/harvester.mp4" background-video-loop=true background-video-muted=true background-opacity=0.4}

1970-1990:

Intensification of Harvesting

Energy & Chemical Pollution Reduction in Pulp and Paper

. . .

2000-2021:

New Materials

## {.center}

::: {.columns}
::: {.column width="33%"}

### [Bioresource]{.text-batlow-darkgreen}

[1970-1990:]{.text-gray-600}

Intensification of Harvesting

Energy Usage & Chemical Pollution Reduction in Pulp and Paper

[2000-2021:]{.text-gray-600}

New Materials

:::
::: {.column width="33%"}

### [Biotechnology]{.text-batlow-forestgreen}

Mostly applications to clean water in pulp and paper

New materials from cellulose

:::
::: {.column width="33%"}

### [Bioecology]{.text-batlow-darkblue}

74% Recycling related

Decommodification of value chains through regional identity products

:::
:::

# Outlook & Aims {.inverted}

## Sharpen Analytical Focus on Directionality

To what extent is there a tension between the normative (desired) and positive (observed) directionality in the transition to a bioeconomy?

How does positive directionality shape normative directionality?

# Take Aways

1. [The bioeconomy is declining]{.text-royal-blue-600} relative to total innovation

2. [Positive directionality towards bioresource vision]{.text-royal-blue-600}, but overlap between visions and open development paths

# References

::: {#refs}
:::