### 2024-11-28 paper.qmd
update website listings

---
title: Quantifying Directionality and Innovation Output in Sweden's Transition to a Forest-Based Bioeconomy
subtitle: Draft
author:
  name: Philipp Jonas Kreutzer
  email: philipp_jonas.kreutzer@ekh.lu.se
  orcid: 0000-0003-4234-9043
  affiliations: 
        - id: LU
          name: Lund University
          department: Department of Economic History
          city: Lund
          country: Sweden
date: last-modified
abstract: |
  The transition to a bioeconomy is high on the agenda for research, industry and policy. Like other sustainability transition, the bioeconomy transition is characterized by contestation over desired outcomes and transition pathways. While shared visions are seen as pivotal in guiding transitions, the bioeconomy discourse is marked by the contention of three competing visions. The Swedish forest-based bioeconomy is well established as a case study in this arena. Yet, knowledge about innovation produced across the entire system is scarce. The use of detailed innovation data from a Swedish literature-based innovation output database, SWINNO, allows to capture the actual innovation output of the Swedish forest-based bioeconomy innovation system. Qualitative coding of the source material links innovation from 1970 to 2021 with visions of a bioresource, biotechnology and / or bioecology bioeconomy. Findings reveal a decrease in the innovation output of the forest-based bioeconomy relative to Sweden's overall commercialized innovation. Notably, innovations associated with a bioresource vision surged during the energy crises of the 1970s, pre-dating their formalization in strategy documents, suggesting a significant role for path-dependency in shaping shared visions. Concurrently, the presence of innovations aligning with multiple visions indicates an open opportunity space for future development across various directions. This research underscores the dynamic interplay between historical contexts, shared visions, and innovation trajectories within the bioeconomy transition, and questions the extent to which visions can provide novel directionality to transitions.
keywords: 
  - Directionality
  - Bioeconomy
  - Forest
  - Transformative Innovation Policy

bibliography:
  - swinno-bioeconomy-directionality.bib

format:
#   pdf:
#     documentclass: scrartcl
#     papersize: a4
#     linestretch: 1.25
#     mainfont: "Adobe Garamond Pro"
#     sansfont: "Adobe Garamond Pro Semibold"
#     indent: true
#     csl: apa.csl
    # link-citations: true
    # link-bibliography: true
    # colorlinks: true
#       
# number-sections: true
# number-depth: 3
#     output-file: "lund_style"
  elsevier-pdf:
    keep-tex: true  
    journal:
        name: Environmental Innovation and Societal Transitions
        formatting: review
        model: 3p
        cite-style: authoryear
        highlights:
            - Highlight 1
            - Highlight 2 
            - Highlight 3

execute:
  warning: false
  error: false
  echo: false
jupyter: python3
---

```{python}
# | label: Imports

import pandas as pd
import numpy as np
import scipy.stats as stats
from IPython.display import display, Markdown
import matplotlib.pyplot as plt
import matplotlib.ticker as mtick
import seaborn as sns
from tabulate import tabulate

from swinno_bioeconomy_directionality.utils import format_table
from swinno_bioeconomy_directionality.data_handling import database
from swinno_bioeconomy_directionality.config import (
    PROJ_ROOT,
    RAW_DATA_DIR,
    PROCESSED_DATA_DIR,
)

conn = database.connect_swinno_db()

sns.set_theme(style="white")
```

```{python}
#| label: Load Data
uncleaned_bioeconomy_query = pd.read_csv(RAW_DATA_DIR / "uncleaned_bioeconomy_query.csv")
bioeconomy = pd.read_csv(RAW_DATA_DIR / "id_year_bioeconomy.csv")
bioeconomy = bioeconomy.loc[bioeconomy["bioeconomy"] == 1]

sni_codes = pd.read_sql_query(
    """
select
*
from 
sni_codes
""",
    conn,
)

sni_codes = sni_codes.rename(columns={"code": "use_sector"})

swinno = pd.read_sql_query(
    """
select sinno_id, year_of_commercialization as year, innovation_name_in_swedish as name
from innovation;
""",
    conn,
)

swinno["bioeconomy"] = (
    swinno["sinno_id"].astype(str).isin(bioeconomy["sinno_id"].astype(str).unique())
)

# full code labels for merging
codes_df = pd.read_sql("SELECT * FROM classification_codes", conn)

codes_df["Category"] = codes_df["Category"].str.strip()
uncertain_eco = {"Code": ["000"], "Category": ["Unsure"]}
codes_df = pd.concat([codes_df, pd.DataFrame(uncertain_eco)], ignore_index=True)

codes_df = codes_df[~codes_df["Code"].duplicated(keep="last")]

codes_df = codes_df.set_index("Code")

eco_innovations = database.query_eco_innovations(conn)
eco_innovations["innovation_type"] = eco_innovations["innovation_type"].str.strip()

bioeconomy_visions = pd.read_csv(RAW_DATA_DIR / "bioeconomy_visions.csv")
```

```{python}
# | label: Innovation Counts
n_swinno = swinno["sinno_id"].nunique()
year_min = swinno["year"].min()
year_max = swinno["year"].max()

n_uncleaned_bioeconomy_innovations = uncleaned_bioeconomy_query["sinno_id"].nunique()
sum_bioeconomy_innovations = int(swinno["bioeconomy"].sum())
```

# Introduction {#sec-introduction}

<!-- Paragraphs 1-2: Motivation. After reading these paragraphs a reader in any field of economics should believe that if you answer your research question your paper will make an important contribution. -->
At the heart of sustainability transitions lies the question how societal functions can be fulfilled within ecological boundaries [@kohler2019AgendaSustainabilityTransitions].
Societal functions are provided through socio-technical systems, which consist of the interplay between the production, distribution, and use of technology [@geels2004SectoralSystemsInnovation].
Transitions are complex, long-term processes of shifts between different systems, involving diverse actors who have to chose between multiple, open-ended pathways, uncertain of their technological, social and political success [@kohler2019AgendaSustainabilityTransitions].
Further, success in the outcome and chosen pathway of sustainability transitions is contested based on the values, norms, and goals of  different actors and social groups.
Hence, innovation cannot be seen as a scalar, but must rather be understood as a vector, as possessing and providing direction to change rather than being simply progress forward [@stirling2011PluralisingProgressIntegrative].
@andersson2021OutcomesDirectionalityMorphology term these two properties _positive_ and _normative_ directionality, respectively.
Positive directionality of innovation arises from the fact that technological change is path-dependent [@arthur1989CompetingTechnologiesIncreasing;@david1985ClioEconomicsQWERTY], while normative directionality is rooted in the assumption that without public policy, private actors lack intrinsic incentives to pursue sustainability [@kohler2019AgendaSustainabilityTransitions].

A sustainability transition which has been gaining relevance for researchers, policy makers, and producers alike is the bioeconomy.
Broadly speaking, it aims to radically shift economic systems from relying on fossil fuels for energy and material resource, to an economic system built on biobased materials [@bennich2018BioBasedEconomyDynamics], circular and renewable value chains [@damato2017GreenCircularBio] and higher environmental integrity through innovation [@bugge2016WhatBioeconomyReview].
As one of the main sources for biomass in a bioeconomy [@priefer2017PathwaysShapeBioeconomy], forests have become a key arena for the bioeconomy transition, highlighting tradeoffs between material provision, carbon storage and biodiversity [@hogbom2021TrilemmaNordicBaltic].
Underneath the umbrella term "bioeconomy", the concept and viable transition pathways are heavily contested [@allain2022BioeconomicsVsBioeconomy].
Three different visions of a bioeconomy compete in discourse to provide normative directionality of the transition: the first vision centers on the provision of biomass and renewable resources, the second on the scientific development and application of biotechnology, while the third emphasizes ecological considerations[@bugge2016WhatBioeconomyReview; @vivien2019HijackingBioeconomy].
Although the discourse is well understood in terms of content [e.g., @bugge2016WhatBioeconomyReview;@vivien2019HijackingBioeconomy] its narratives [@bauer2018NarrativesBiorefineryInnovation], and actors [@holmgren2022WhoseTransformationThis], the material effects of this transition, however, remain important research topics [@fischer2020SwedishForestsBioeconomy].
How many innovations are commercialized in this new economy, especially considering the central role of innovation in bioeconomy discourse? And what does commercialized innovation suggest for the positive directionality of the bioeconomy transition?

<!-- @fischer2020SwedishForestsBioeconomy also note the consensus builidng of Swedish bioeconomy discourse marginalizes non-mainstream opinions. I can show that this is rooted, at least in part, in the path-dependence on technology output. -->

<!-- • Paragraphs 3-4: Challenges. These paragraphs explain why your research question has not already been answered, i.e., what are the central challenges a researcher must tackle to answer this question. -->
<!-- • Paragraph 5: This Paper. This paragraph states in a nutshell what the paper accomplishes and how. -->
<!-- • Paragraphs 6-7: Model. Summarize the key formal assumptions you will maintain in your analysis. -->
<!-- • Paragraphs 8-9: Data. Explain where you obtain your data and how you measure the concepts that are central to your study. -->
<!-- • Paragraphs 10-11: Methods. Explain how you take your model to the data and how you overcome the challenges you raised in paragraphs 3-4. -->

To unpack the plurality of possible socio-technical development paths, requires high data quality when measuring innovation.
Although progress towards monitoring the bioeconomy transition has been made [@jander2019MonitoringTransitionBioeconomy; @jander2020MonitoringBioeconomyTransitions], assessing directionality is limited by a lack of innovation output measures [@wydra2020MeasuringInnovationBioeconomy].
In this paper, I use SWINNO, a literature-based innovation output database capturing all significant commercialized innovation in Sweden [@sjoo2014DatabaseSwedishInnovations], to analyze the innovation output of the Swedish forest-based bioeconomy.
I develop and apply a classification scheme to identify which bioeconomy vision an innovation produced or used in forestry or the downstream value chain aligns with.

<!-- • Paragraphs 12-13: Findings. Describe the key findings. Make sure they connect clearly to the motiva-
tion in paragraphs 1-2. -->
Results show that Swedish bioeconomy innovation output has declined over the last 50 years.
Of the three different visions the bioresource vision is the most pronounced, in line with expectations from discourse analysis and the Swedish bioeconomy strategy.
However, the strategy should not be seen as visionary, as the innovation system had been producing bioeconomy innovation and even bioresource aligned innovation from the 1970s onwards.
This points to a strong interconnection between path-dependence and the creation of directionality in the policy space.

<!-- • Paragraphs 14-15: Literature. Lay out the two main ways your paper contributes to the literature. Each paragraph should center around one contribution and should explain precisely how your paper differs from the most closely related recent work. -->

<!-- By analyzing the innovations produced by the forest-bioeconomy system in Sweden, this paper adds to existing literature on directionality in sustainability transitions and bioeconomy literature in two main ways.
On one hand, using literature covering innovation output can be used to assess the directionality a sociotechnical system  -->

<!-- From an empirical perspective, this is essentially about furthering the current understanding of how sociotechnical systems grow as a result of historical path-dependencies, system-internal dynamics and external influence, towards a more elaborate understanding of how these mechanisms influence how system configurations change over time. From a conceptual perspective, it is rather about enabling such investigations, by developing analytical frameworks that capture not only the pace, but also the directionality of innovation processes. For example, there may be a need to advancethe commonly used technological innovation systems framework, since the typology of functional processes is geared towards investigating system growth in a way that hides the potential for vastly different technical, social and spatial development trajectories (Andersson, 2020).
@andersson2021OutcomesDirectionalityMorphology [127-128] -->

This article has a dual purpose.
On one hand, it examines the directionality of Sweden's bioeconomy transition using innovation output data matched to public funding applications.
On the other hand, it examines the idea that visions are suitable to steer transitions in normatively desired directions.
First it finds that only about half of significant commercialized innovation in Sweden corresponds to any bioeconomy vision at all, while the remainder can better be described as general improvements to exisiting technology.
A point here that vision have possibly only recently been formulated coherently enough to develop guiding power.
In addition, despite the bioresource vision dominating discourse the bioeconomy transition is still highly contested.
Comparing innovations aligned with each vision shows that the bioresource vision does indeed account for a majority of vision aligned innovations.
However, by extending the temporal horizon into the last century this paper further shows that innovation alinged with the bioresource vision has dominated other visions since the 1970s, therefore casting doubt on the external directionality of visions formulated at the policy level relative to an endogenous vision shared by key / a majority of innovation producers.
For the Swedish case study this raises the question of who is horse and who is carrot in drafting and concentrating on a vision for normative directionality.
For transitions at large this raises additional questions over the effectiveness of visions to provide directionality to transitions, such as to which extent do existing technology sets influence the formulation of vision?
Can visions for transforming existing socio-technical systems be formulated without the buy-in of regime actors?

<!-- TODO: Read up on proponents of bottom up vision generation  -->

# Literature Review

## Transformative Innovation Policy

### Directionality

#### The Role of Visions

## Contested Transition Towards Bioeconomy in Sweden

### Differences in Visions

###

# Data and Methods {#sec-data-methods}

Operationalizing directionality suffers from frequent lack of differentiation of positive and normative directionality, in addition to inherent issues in the measurement of innovation.
Notable exceptions are the work on socio-technical system morphology by @andersson2021OutcomesDirectionalityMorphology, who identify changes in the configuration of three different renewable energy systems, and by @yang2022ShapingDirectionalitySustainability who use interviews and secondary source analysis to understand changes to insitutions in solar photovoltaic production in two Chinese provinces.
However, both studies deal more with the aggregate effects of directionality on socio-technical systems rather than with positive directionality of innovation directly.

Normative directionality can and has been assessed through analysis of discourse [@] and narratives [@bauer2018NarrativesBiorefineryInnovation], visions and aims expressed in strategy documents [@parks2022DirectionalityTransformativeInnovation], and R&D expenditure [@cappellano2022WhereInnovationMeets; @wydra2020MeasuringInnovationBioeconomy].
Where positive directionality of technological change has been assessed, scientific publications [@wydra2020MeasuringInnovationBioeconomy; @jankovsky2021InnovationsForestBioeconomy], and more often, patents are used [e.g., @aghion2016CarbonTaxesPath;@popp2005LessonsPatentsUsing;@moscona2022DoesDirectedInnovation].
Such data, however, suffer from limitations when attempting to assess the directionality of socio-technical systems.
While measures of normative directionality are equipped to identify the desired direction by different actors, visions need to first be accepted by innovators and reformulated into search.
R&D funding can help in this task, but data on R&D funding is often too broad, targeting entire sectors or specific industries [@popp2005LessonsPatentsUsing].
While patents are more granular and a closer proxy of technological change, they suffer from well documented measurement errors [@popp2005LessonsPatentsUsing]; not all innovations are patented, and not all patents become innovations [@arundel1998WhatPercentageInnovations].

Literature-based innovation output measures remedy these issues, by capturing commercialized innovations [@vanderpanne2007IssuesMeasuringInnovation].
Applying this method to  editor-reviewed articles published in 15 independent trade journals covering technological developments for general and highly specialized audiences, the SWINNO database contains `{python} n_swinno` innovations commercialized by Swedish companies between `{python} year_min` and `{python} year_max` [@sjoo2014DatabaseSwedishInnovations].
Since this method relies on the judgment of editiors and journalists, it can be described as significant innovations, omitting small reconfigurations of exisiting products and processes.
The SWINNO database has significant benefits compared to other innovation measures.
For one, it ensures that actual innovation, meaning commercialized products and processes, and thus that its positive directionality is captured.
Second, because the source articles are preserved, it provides a rich source of data, frequently providing insights into the aims, motivations, advantages and challenges of the covered innovation, in addition to descriptions of what the innovation consists of, which allows to assess and quantify the direction of technological change.

To capture the forest-based bioeconomy the database was searched for innovation associated with core forest sectors and keywords describing forest products which may be used in sectors along the bioeconomy value chain based on @wolfslehner2016ForestBioeconomyNew.
The list of keywords and sectors is presented in the appendix (@sec-appendix-data), additionally reproducible code to generate the figures and tables in this paper can be found at [`github.com/pjkreutzer/swinno-bioeconomy-directionality`](https://github.com/pjkreutzer/swinno-bioeconomy-directionality).

## Classification Methodology {#sec-classification_method}

I carefully read the source articles of innovations identified in the database and tagged their associated unique identifiers with two types of labels.
The first label is the main interest of this study: the bioeconomy vision an innovation can be associated with.
The second label provides a deeper understanding if and how the innovation reduces environmental impacts; not all bioeconomy innovations need to be sustainability oriented innovation [@broring2020InnovationTypesBioeconomy].

### Bioeconomy Visions {#sec-bioeconomy_classification_method}

Based on the analyses of competing bioeconomy visions in @bugge2016WhatBioeconomyReview and @vivien2019HijackingBioeconomy, I developed and applied a matrix of characterstics for each vision.
While the criteria for the bioresource and bioecology vision are entirely based on the two previous publications, the biotechnology criteria are additionally based on a literature review of @yeung2019CurrentResearchBiotechnology, concerning biotechnology publications.

Reading the source texts of innovations identified from SWINNO, I classified an innovation as being *aligned* with a vision if it fulfills at least one criteria in the matrix.
An innovation can align with multiple visions, for example, where biomass is converted and upgraded (bioresource vision) by application of biotechnology (biotechnology vision).
Or the value chain decommodifying introduction of new production lines by a family-business which use wood from their own forests to produce houses with high territorial identity (*Svensk trävaru- & pappersmassetidning*, 1976:7, pp. 539,541,544), that aligns with the bioresource and bioecology vision.
Lastly, there is also the possibility that an innovation cannot be placed into at least one of the visions.
This is the case for innovation which brings general technological improvements, purely advances economic performance, or relates to a general purpose technology such as appliance of computers.
These innovations are labeled as Vision Neutral, which does not imply that the innovation is not used, or originates in the forest based bioeconomy as defined above.
An illustrative example of such an innovation are improved work boots (*Sågverken*, 1974:2, p. 121), which promote safety of forest workers but do not correspond with any of the criteria in @tbl-decision-matrix.

| Corresponding Vision | Source article explicitly mentions that innovation:                                                   |
|----------------------|-------------------------------------------------------------------------------------------------------|
| Biotechnology        | Develops new, or modifies existing, organism, biological system, or process                            |
|                      | Uses cellular or molecular process to solve problem or in production process                          |
|                      | Applies biological agent in scientific or engineering based principle to produce product or service   |
|                      | Based on molecular biology, biochemistry, cell biology, embryology, genetics or microbiology research |
|                      | Provides method for research in the above areas                                                       |
|||
| Bioresource          | Biomass used to substitute other input                                                                |
|                      | Refines biomass through mechanical, chemical or other process                                         |
|                      | Converts and upgrades bioresource                                                                     |
|                      | Optimises land use                                                                                    |
|                      | Includes degraded land in production of biofuels                                                      |
|                      | Intensifies use of biomass, including from waste                                                      |
|||
| Bioecology           | Prioritizes biodiversity, conservation, ecosystem or soil health                                      |
|                      | Integrates production systems                                                                         |
|                      | Develops high quality product with territorial identity                                               |
|                      | Reuses or recycles waste                                                                              |
|                      | States compatibility with biosphere as motivation or aim                                              |
|                      | States need for social or ecological counter-expertise relative to other socio-technical innovations  |
|                      | Promotes or uses a circular production mode                                                           |
|                      | Promotes or uses a self-sustained production mode                                                     |

: Decision matrix used to classify innovation. {#tbl-decision-matrix tbl-colwidths="[20,80]"}

# Results {#sec-results}

```{python}
#| label: Calculate Bioeconomy Share

def calculate_bioeconomy_share(swinno):
    df = swinno.copy()
    df = df[["sinno_id", "year", "bioeconomy"]]
    df = (
        df.groupby("year")
        .agg(
            total_count=("sinno_id", "nunique"), bioeconomy_count=("bioeconomy", "sum")
        )
        .reset_index()
    )
    df["bioeconomy_share"] = df["bioeconomy_count"] / df["total_count"]

    return df

bioeconomy_share = calculate_bioeconomy_share(swinno)
```

```{python}
# | label: Bioeconomy Trend Description
# | output: asis
share_max = bioeconomy_share["bioeconomy_share"].max()
share_max_year = bioeconomy_share.loc[
    bioeconomy_share["bioeconomy_share"] == share_max, "year"
].values[0]
share_last_year = bioeconomy_share["year"].max()
share_latest = bioeconomy_share.loc[
    bioeconomy_share["year"] == share_last_year, "bioeconomy_share"
].values[0]

pct_diff_max_latest = (share_max - share_latest) * 100
total_max = bioeconomy_share["total_count"].max()
total_max_year = bioeconomy_share.loc[
    bioeconomy_share["total_count"] == total_max, "year"
].values[0]

bioeconomy_max = bioeconomy_share["bioeconomy_count"].max()
bioeconomy_max_year = bioeconomy_share.loc[
    bioeconomy_share["bioeconomy_count"] == bioeconomy_max, "year"
].values[0]

total_bioeconomy_share = sum_bioeconomy_innovations / n_swinno

```

Querying the SWINNO database for forest based bioeconomy innovations initially yields `{python} n_uncleaned_bioeconomy_innovations` innovations. False positives were removed from the dataset during classification, resulting in `{python} sum_bioeconomy_innovations` innovations which originate from, are used in, or make use of products derived from the forest.

## Innovation Output of the Swedish Forest-Based Bioeconomy {#sec-results-innovative}

In total, the forest-based bioeconomy produced `{python} f"{total_bioeconomy_share:.0%}"` of innovation captured in the SWINNO database.
While the total innovation output and bioeconomy innovations feature a period of high output from the 1970s to the 1990s, with a peak in `{python} total_max_year` for total output and `{python} bioeconomy_max_year` for the bioeconomy, the economic downturn and industrial restructurization of the 1990s was accompanied with a decline in commercialized innovation.
Unlike the other sectors and socio-technical systems, however, it appears that the bioeconomy was not able to recover to the same extent.
Its percentage of total innovation declined by `{python} f"{pct_diff_max_latest:.0f}"` percentage points from an all-time high of `{python} f"{share_max:.0%}"` in `{python} share_max_year` to `{python} f"{share_latest:.0%}"` in `{python} share_last_year`. 
@fig-innovation-trends illustrates this using the five year moving average for the number of innovations commercialized per year in the bioeconomy and in the entire dataset, including the bioeconomy.
The decline in relative importance of bioeconomy innovations can be identified in the share of bioeconomy innovation of total innovation.

![a. A total number of innovation registered in SWINNO database. b. Forest-based bioeconomy innovations registered in SWINNO database. c. Percentage of forest based bioeconomy innovations to total innovation registered in SWINNO database. Shown are the 5 year moving averages for each time series.](results/figures/share_test.tex){#fig-innovation-trends width=100%}

##  Visions of the Swedish Forest Bioeconomy Innovation System

```{python}
#| label: Uncertain Visions

bioeconomy_visions.loc[
    bioeconomy_visions["bioeconomy_vision"] == "Not Bioeconomy",
    ["bioeconomy_vision"],
] = "Vision Neutral"

uncertain_bioeconomy_visions = bioeconomy_visions.query("bioeconomy_vision == 'Unsure'")

certain_bioeconomy_visions = bioeconomy_visions.loc[
    ~bioeconomy_visions["sinno_id"].isin(uncertain_bioeconomy_visions["sinno_id"]), :
]

count_certain_bioeconomy_visions = (
    certain_bioeconomy_visions.groupby(["year", "bioeconomy_vision"])
    .size()
    .reset_index(name="count")
)

total_certain_bioeconomy_visions = (
    count_certain_bioeconomy_visions.groupby("bioeconomy_vision")
    .agg("sum", numeric_only=True)
    .sort_values("count", ascending=False)
    .reset_index()
)
```

```{python}
#| label: Number of Uncertains
#| output: asis

n_uncertain_bioeconomy_visions = uncertain_bioeconomy_visions["sinno_id"].nunique()
```

```{python}
# | label: Number of Results
# | output: asis

n_certain_bioeconomy_visions = certain_bioeconomy_visions["sinno_id"].nunique()

certain_bioeconomy_visions_min_year = certain_bioeconomy_visions["year"].min().item()

certain_bioeconomy_visions_max_year = certain_bioeconomy_visions["year"].max().item()

n_most_bioeconomy_vision = total_certain_bioeconomy_visions.iloc[0, 2].item()

def calculate_share(a: str):
    return (
        total_certain_bioeconomy_visions.loc[
            total_certain_bioeconomy_visions["bioeconomy_vision"] == a, "count"
        ].values[0]
        / n_certain_bioeconomy_visions
    )

share_vision_neutral = calculate_share("Vision Neutral")
share_bioecology = calculate_share("Bioecology Vision")
share_bioresource = calculate_share("Bioresource Vision")
share_biotechnology = calculate_share("Biotechnology Vision")

vision_1_count = total_certain_bioeconomy_visions.iloc[0, 2].item()
vision_2_count = total_certain_bioeconomy_visions.iloc[1, 2].item()
vision_3_count = total_certain_bioeconomy_visions.iloc[2, 2].item()
vision_4_count = total_certain_bioeconomy_visions.iloc[3, 2].item()

vision_1_count = total_certain_bioeconomy_visions.iloc[0, 2].item()
vision_2_count = total_certain_bioeconomy_visions.iloc[1, 2].item()
vision_3_count = total_certain_bioeconomy_visions.iloc[2, 2].item()
vision_4_count = total_certain_bioeconomy_visions.iloc[3, 2].item()

vision_1 = total_certain_bioeconomy_visions.iloc[0, 0]
vision_2 = total_certain_bioeconomy_visions.iloc[1, 0]
vision_3 = total_certain_bioeconomy_visions.iloc[2, 0]
vision_4 = total_certain_bioeconomy_visions.iloc[3, 0]

n_bioresource_1970_1990 = certain_bioeconomy_visions.query(
    "year.between(1970, 1990) & bioeconomy_vision == 'Bioresource Vision'"
)["sinno_id"].nunique()
```

```{python}
# | label: Count Recycling Innovation

recycle_innovation = pd.read_sql(
    """
SELECT *
FROM classification_codes
LEFT JOIN eco_innovations ON classification_codes.Code = eco_innovations.innovation_type
WHERE eco_innovations.innovation_type = 107 OR eco_innovations.innovation_type = 206 AND sinno_id NOT IN (SELECT sinno_id FROM categorization_notes WHERE notes LIKE '%not forest%')
""",
    conn,
)

recycle_innovation_count = recycle_innovation["sinno_id"].nunique()

n_bioecology_vision = count_certain_bioeconomy_visions.loc[
    count_certain_bioeconomy_visions["bioeconomy_vision"] == "Bioecology Vision",
    "count",
].sum()
```

For the following part, `{python} n_uncertain_bioeconomy_visions` innovations whose source text did not allow labeling without significant uncertainty are excluded. These innovations are planned to undergo a revision by subject matter experts at a later point.

Among the remaining `{python} n_certain_bioeconomy_visions` innovations,
`{python} total_certain_bioeconomy_visions.iloc[0, 2]` (`{python} f"{share_vision_neutral:.0%}"` of all commercialized innovation) were not aligned with any vision. Next the `{python} vision_2.lower()` was associated with `{python} vision_2_count` (`{python} f"{calculate_share(vision_2):.0%}"`), while `{python} vision_3_count` (`{python} f"{calculate_share(vision_3):.0%}"`) and  `{python} vision_4_count` (`{python} f"{calculate_share(vision_4):.0%}"`) innovations were classified as `{python} vision_3.lower()` and `{python} vision_4.lower()`, respectively.
Note that the sum of all categories does not equal the total number of classified innovations, as one innovation can be associated with more than one bioeconomy vision.
@fig-bioeconomy-trends shows the count of innovation for each categorization per year.
Output for the biotechnology and bioecology vision appears inconsistent, with years in which not a single commercialized innovation aligned with these visions.
Additionally, due to the low number of innovations aligned with these visions no clear trends can be identified.
In contrast, the two peaks of total innovation output of @fig-innovation-trends a, can be identified in the bioresource and vision neutral innovations.
Unlike vision neutral innovation, bioresource innovation declined to under 4 commercialized innovations after 2010.

![Counts of innovation classification by Bioeconomy Vision Category. Innovations can be classified into multiple visions simulatenously. The sum of all vision counts does not match the sum of all bioeconomy innovations.](results/figures/vision_plot.tex){#fig-bioeconomy-trends width=100%}

### A Deeper Look at Innovations in Each Vision {#sec-results-deeper-visions}

#### Bioresource Vision

Although the bioresource vision claims the most aligned innovations, its most active period in terms of commercialized innovations was at the beginning of this study's time frame. `{python} n_bioresource_1970_1990` innovations were commercialized between 1970 and 1990.
These can be grouped into two themes.

The first theme addressed environmental issues, especially greening the pulp and paper production process and has been extensively documented, for example by Bergquist and Söderholm [-@bergquist2011GreenInnovationSystems; -@bergquist2015TransitionGreenerPulp], and @bergquist2018GreeningPulpPaper.
The second innovation group was a response to the energy crisis resulting from oil export restrictions placed by OPEC countries in the 1970s [@bergquist2016SustainableEnergyTransition].
Once again a major player was the pulp and paper industry which commercialized numerous innovations to improve energy efficiency in drying (*Ny Teknik*, 1972:1, p. 4), reduce cool down of furnaces (*Kemisk Tidskrift*, 1976:9 pp. 96,99-100) and generate cheaper energy from waste, byproducts (*VVS*, 1980:7-8, p.69) and excess process heat (*Kemisk Tidskrift*, 1981:7, p. 82).
The second major player was the forestry sector.
A cluster of technologies such as forwarders, tree harvesters, wood driers and processors emerged in response to a wood shortage and soaring energy prices from the oil shocks in the 1970s [@taalbi2014InnovationCreativeResponse].
Key innovations in this technology cluster were the pendulum saw (*Sågverken*, 1977:2, pp. 97-101,125), which could be mounted on vehicles to mechanically fell entire trees at once (*Sågverken*, 1974:10, pp. 725-727), and vehicles that, despite being heavy, limited the damage to soft forest soil and importantly, the roots therein (e.g., *Sågverken*, 1974:10, p. 725-727; *Sågverken*, 1975:10, p. 731).

The second period of increased activity starting in 2000 focused more on
upgrading of biomass into new products such as construction timber (e.g., *NTT* 2003:27, p. 11; *NTT*, 2003:8, p.6; *NTT* 2004:18, pp.16-17), various paper products (e.g., *Kemivärlden Biotech med Kemisk Tidskrift*, 2009:11, pp. 60-61; *SPCI Svensk Papperstidning*, 2005:7, pp.10-11), but also biobased alternatives to plastics (e.g., *SPCI Svensk Papperstidning*, 2018:4, pp. 28-33; *SPCI Svensk Papperstidning*, 2019:6, p.24; *Plastforum*,2019:5, p.31), or refined bio-fuel alternatives to fossil fuels, such as the LignoBoost Process (*SPCI Svensk Papperstidning*, 2006:7, pp. 14-16), or gassification of black liquour -- a by-product of pulp digestion -- for electricity generation (*Ny Teknik*, 1990:16, p.5) .
A second theme is rationalization of production processes, especially through the application of computers and software (e.g.,  *NTT*, 2007:14, p. 13; *Automation*, 2021:1, p. 24).

Therefore, while the first period of high innovation output until the 1990s was characterized by attempts to modernize timber production, mainly through the mechanization of harvesting, and cleaning the pulp and paper process, the second period of active commercialization focused on finding new applications for biomass and improving the cost effectivness of exisiting processes.
This is illustrated with innovations in forwarders -- specialized machines to aid in the removal of timber from harvest sides before loading them onto transport -- where of 9  innovations, 7 were commercialized before 1990.
Moreover, the two innovations commercialized after focused on greening fuel consumption through hybrid engines (*Ny Teknik*, 2008:34, p. 24) and weight reduction (*Ny Teknik*, 2008:34, p. 24; *Automation*, 2012:40, p. 26).
Meanwhile, earlier innovations focused on establishing the product category by solving issues of soil damage from use of heavy machinery (*Sågverken*,1975:10, p. 731; *Sågverken*, 1982:10, p. 77) as well as integrating forwarders with other technologies for planting (*Sågverken*, 1979:3, p. 237) and removing trees (*Ny Teknik*, 1976:43, p.20; *Sågverken*, 1975:12, pp. 905,907).
<!-- TODO make dynamic -->

```{python}
#| label: forwarder
#| include: false
forwarder_innovation = pd.read_sql(
    "SELECT sinno_id, year_of_commercialization from innovation where description_in_swedish LIKE '%skotare%'",
    conn,
)

forwarder_innovation
```

#### Biotechnology Vision

Only 37 innovations were aligned with the biotechnology vision.
Of those, the majority focused on cleaning waste water from, for example, fibers (e.g., *Kemisk Tidskrift*, 1972:4, p. 19), or chemicals (e.g., *Ny Teknik*, 1970:4, p. 8), with applications mostly in the pulp and paper industry.
Other innovation applied biomass derived from the forest in new products, such as new fibers (*Ny Teknik*, 1976:9, pp. 1,4-5), or applications of cellulose in heat isolation materials (*Ny Teknik*, 2002:9, p. 14).
Innovations in this vision frequently overlapped with bioresource innovations, such as the above mentioned LignoBoost process.

#### Bioecology Vision

Recycling accounted for `{python} recycle_innovation_count` unique innovations, or `{python} f"{(recycle_innovation_count / n_bioecology_vision):.0%}"` of innovation associated with the bioecology vision.
The majority of innovations in aligned with this vision due to some explicitly stated environmental aim within existing production methods.
An example is an instrument that measures the extent to which a tree is decayed from the inside (*NTT*, 2006:2, p. 20).
While such trees are of little economic value, they play a big role in biodiversity conservation if left standing.
This innovation aims to identify these trees before they are felled and thus to protect it from being felled.
Some innovations stood out as attempts to generate social or ecological counter expertise.
For example, in 1971 *Sågverken* (3, p. 195), covered an innovation in wooden floors made from wide, red pine planks.
Notable about this innovation was that it countered the trend of economic rationalization of the time, in favor of tight integration between production and forest ownership.
Another example of regional timber being treated not as a commodity, but as a desirable high quality product is the Masonite sytem for pre-fabricated wooden house (*Svensk trävaru- & pappersmassetidning*, 1976:7, pp. 539,541,544).
The specifc growing conditions of Northern Sweden are provided as a reason for denser wood, especially well suited for construction due to its strength properties and the "[t]otal integration makes the company feel optimistic: own forests, top-modern sawmill, board mill, newly built component factory are positive factors meaning advantages in all stages" (p. 544).
Similiar decommodification of value chains in favor of local product could even be found in the paper industry, where fresh wood from within a 100km radius was found to produce better quality and reduce chemical and energy usage (*SPCI Svensk Papperstidning*, 2015:8, pp. 26-30).

# Discussion {#sec-discussion}

This discussion first turns towards the directionality of the Swedish forest bioeconomy before extracting insights from the case on the interplay between visions and directionality in sustainability transitions.

## Positive Directionality of the Swedish Forest-Based Bioeconomy and Possible Future Transition Paths

The decline in innovations commercialized in the forest-based bioeconomy relative to all commercialized innovations suggests that Sweden's capacity to produce technological change towards a bioeconomy has diminished over the last 50 years.
While innovation counts should be interpreted with caution, as the magnitude of technological change stemming from individual innovations can differ, it is surprising that a economic sector, which is expected to realize ambitious economic, social and ecological targets, is losing importance in relation to other Swedish innovations.
While it appears that the crises in forestry and energy of the 1970s stimulated innovation directed towards economic rationalization of forestry, the bioeconomy could not benefit from similar stimulants when exploring conversion of biomass into new products after the financial crisis in 2008, and the shock to timber production after storm Gudrun in 2005.

As for the positive directionality of the Swedish transition to a forest-bioeconomy, the dominant vision appears to be successful in so far as it is the vision most innovations align with.
That a majority of innovation appear neutral in regards to vision based directionality is not to say that these do not posess positive directionality.
These innovations should be regarded as general improvements to exisiting technology and hence are not neutral in regards to technological change, only in regards to not aligning with normatively directionality.
The congruency between positive and normative directionality can be explained by a specific feature of Swedish socio-technical systems: a strong desire to create consensus [@belandlindahl2015SwedishForestryModel].
@bergquist2011GreenInnovationSystems illuminate how strong collaboration networks between public and private actors shaped the greening of Sweden's pulp and paper industry between 1960 and 1980.
The findings of this study support this view.
Rather than providing directionality towards sustainability, preparations for a national forest strategy and the national forest program strongly align with the trajectories of technological change produced by the innovation system.
While innovation systems literature highlights this as the mark of well implemented transition management, this consensus building also comes at a cost to transition paths.
However, this consensus building comes at a cost, where dissenting views are marginalized [@fischer2020SwedishForestsBioeconomy].
@holmgren2022WhoseTransformationThis show that industry interest dominate the transformation to a bioeconomy.
@fischer2020SwedishForestsBioeconomy even go so far to liken it to consensus building through exclusion.

Looking at the direction embedded in commercialized innovations suggests are more nuanced picture.
While the bioresource vision is the most pronounced out of the three visions for a bioeconomy, innovations can and do align with multiple visions at once.
In addition, those innovations that align with any of the bioeconomy visions are more often eco-innovations than vision neutral technological change.
However, it must be said that whether this generates positive environmental impact depends on a number of factors such as technological diffusion and possible rebound effects.

## Visions as Instruments of

While this paper cannot proof conclusively that the visions of future bioeconomy pathways are reformulations of available technologies, the innovation output does suggest that at least in this case the vision dominating bioeconomy transition discours is also the vision most innovation output aligns with.
Considering that @holmgren2022WhoseTransformationThis showed that the discourse is captured by

The fact that innovations aligned with the bioresource vision have been commercialized since the

## Limitations {#sec-limitations}

In regards to the above, an important limitation is the use of innovation as a measure of socio-technical transitions.
While the results of this study agree with the survey results of 
@lovric2020FactorsDevelopmentInnovations and capture the production of technological change, it does not directly capture its adoption and use.
In addition literature-based innovation output measures have data limitations might result in a bias against non-commercialized innovation, especially for research into forest sector innovation where organizational innovations such as silviculture management can be important levers to improve, for example biodiversity or cultural ecosystem services.
Such new processes, or new organizational approaches, will only be captured if they are sold to another company.
Relevant product innovation which facilitates these plans, for example products related to the planting of trees, on the other hand, are captured.
A likely consequence of this limitation is that the number of innovation related to the bioecology vision could be under counted, as this vision highlights "[c]ounter-expertise rather than concrete technical solutions" [@vivien2019HijackingBioeconomy 194].

A second limitation arising from the data is the usage of innovation counts.
While the literature-based innovation output captures significant innovation rather than mere reconfiguration, counting innovation thus captured can be misleading as to the impact of each individual innovation.
Some technologies require few break-through innovations before meeting large market demand and diffusing quickly.
Others might go through multiple rounds of more incremental improvements.
Although literature-based innovation output measures assume editors at trade journales already discard incremental changes to include only important improvements, they cannot ex-ante anticipate which innovation will be most important in terms of impact [@kander2019InnovationTrendsIndustrial].
Therefore the decline in innovation output this paper identifies should not be taken to necessarily imply a decline in the importance of the commercialized innovation.
Similarily, the positive directionality measured in this paper should be seen more as a the vector of change embedded in the transition to a bioeconomy, rather than a scalar of its progress.

# Policy Implications {#sec-policy}

Some recommendations that follow from this study directly for Swedish policy makers include:

1. To align innovation output with key aspects of the Swedish vision for a forest-based bioeconomy, there needs to be more focus on high-tech conversion of biomass to high-value added products including, but not limited to, biorefinery.
2. Technological innovation whitin forestry has mainly focused on mechanization of harvesting. More proactive policy interventions can help spur innovation facilitating other silvioculture practices.

# Conclusion {#sec-conclusion}

In conclusion, the technological artefacts produced in the transition to a forest-based bioeconomy in Sweden show a clear directionality aligned with the bioresource vision, which is most favored by central actors and policy makers.
At the same time, there is evidence that the innovation system has produced knowledge aligned with alternative transition pathways towards an ecological bioeconomy, such as decommodification of value chains as well as efforts to increase the circularity of the system and to reduce environmental pollution.
The temporal distribution of innovation output aligns strongly with the Swedish tendency to produce consens oriented policy by aligning policy objectives with existing technolgical trajectories.
While this approach is pragmatic and has produced good outcomes in the past, it is at least doubtful how radical a shift can be expected for the bioeconomy transition, considering that innovation output since the 1970s has not meaningfully explored alternatives to the intensification of biomass resource extraction.

{{< pagebreak >}}
# References {.unnumbered}

:::{#refs}
:::

## Trade Journals

\noindent
*Automation* (1973-2007).
Sundbyberg: Verkstadstidningen Förlags AB

\noindent
*Kemisk tidskrift: svensk kemisk tidskrift, kemiteknik.* (1969-1993). Stockholm: Ingenjörsförlaget.

\noindent
*Kemivärlden, Biotech, Kemisk tidskrift.* (2005-2018) Stockholm : Mentor Communications

\noindent
*Ny teknik.* (1970-2024). Stockholm: Teknisk tidskrifts förlag.

\noindent
*NTT såg & trä: Nordisk träteknik, Sågverken, Träindustrin.* (2003-2024).
Lidingö: Mentor Communications

\noindent
*Svensk papperstidning: medlemstidning för Svenska pappers- och cellulosaingeniörsföreningen - SPCI.*(2003-2024). Stockholm: AB Svensk
papperstidning.

\noindent
*Svensk trävaru- & pappersmassetidning. *(1970-1990). Stockholm: AB
Svensk trävarutidning.

\noindent
*Sågverken: tidskrift för sågverksindustri, trävaruhandel och skogsbruk.* (1974 - 1998). Stockholm: Svensk trävarutidning.

\noindent
*VVS: tidskrift för värme-, ventilations- och sanitetsteknik : organ för Sven-
ska värme- och sanitetstekniska föreningen.* (1970-1982). Stockholm:
VVS.

# Appendix {.appendix}

<!-- Definition An ecological economy, that is compatible with the biosphere A science-based economy driven by industrial biotechnology A biomass-based economy Nature/economy relations Struggle against entropy and coevolution with the biosphere. Economic development in line with biological evolution The cell is a factory Technology has the power to “correct God's mistakes” Biomass replaces fossil fuels and mining to produce energy and materials Science, technology and society Megatrends with “promethean technologies” Pending the advent of Prometheus III, the economics of prudence and sharing Prospect of a fifth Kondratiev wave based on the rise of biotechnology. The “economy of techno-scientific promises” Biorefining at the heart of ecological transition (multilevel perspective). The economy of learning Sustainability “Strong sustainability” approach and degrowth perspective Very “weak sustainability” approach “Weak sustainability” approach -->

## Data Appendix {#sec-appendix-data}
```{python}
# | label: tbl-full-codes
# | tbl-cap: List of Eco-Innovation Labels
format_table(codes_df.iloc[:5, :])
```

{{< include _appendix-data.qmd >}}