### 2025-04-25 _code_setup.qmd

```{python}
# | label: Imports

import polars as pl
import numpy as np
import scipy.stats as stats
from IPython.display import display, Markdown
import matplotlib.pyplot as plt
import matplotlib.ticker as mtick
import seaborn as sns
from tabulate import tabulate

from swinno_bioeconomy_directionality.utils import format_table

from swinno_bioeconomy_directionality.data_handling.io import DataLoader
from swinno_bioeconomy_directionality.data_handling import database
from swinno_bioeconomy_directionality.config import (
    PROJ_ROOT,
    RAW_DATA_DIR,
    PROCESSED_DATA_DIR,
)

conn = database.connect_swinno_db()

sns.set_theme(style="white")
```

```{python}
# | label: Load Data
from_raw = DataLoader(RAW_DATA_DIR)
from_processed = DataLoader(PROCESSED_DATA_DIR)

uncleaned_bioeconomy_query = pl.read_csv(
    RAW_DATA_DIR / "uncleaned_bioeconomy_query.csv", ignore_errors=True
)

id_year_bioeconomy = from_raw.load_id_year_bioeconomy()

# bioeconomy_ids = id_year_bioeconomy.filter(pl.col("bioeconomy") ==  "1")

swinno = from_raw.load_swinno()

bioeconomy_visions = from_processed.load_visions()
```

```{python}
# | label: Innovation Counts
n_swinno = swinno.select("sinno_id").n_unique()
year_min = swinno["year"].min()
year_max = swinno["year"].max()

n_uncleaned_bioeconomy_innovations = uncleaned_bioeconomy_query["sinno_id"].n_unique()
sum_bioeconomy_innovations = int(id_year_bioeconomy["bioeconomy"].sum())
```

```{python}
# | label: Calculate Bioeconomy Share

def calculate_pct_bioeconomy(swinno):
    return (
        id_year_bioeconomy.group_by("year")
        .agg(
            pl.col("sinno_id").n_unique().alias("n_innovation"),
            pl.col("bioeconomy").sum().alias("n_bioeconomy"),
        )
        .with_columns(pct_bioeconomy=pl.col("n_bioeconomy") / pl.col("n_innovation"))
    ).sort("year")

pct_bioeconomy = calculate_pct_bioeconomy(swinno)
```

```{python}
# | label: Bioeconomy Trend Description
# | output: asis
pct_max = pct_bioeconomy.select("pct_bioeconomy").max().item()
pct_max_year = (
    pct_bioeconomy.filter(pl.col("pct_bioeconomy") == pct_max).select("year").item()
)
pct_last_year = pct_bioeconomy.select("year").max().item()
pct_latest = (
    pct_bioeconomy.filter(pl.col("year") == pct_last_year)
    .select("pct_bioeconomy")
    .item()
)
pct_diff_max_latest = (pct_max - pct_latest) * 100

total_max = pct_bioeconomy.select("n_innovation").max().item()
total_max_year = (
    pct_bioeconomy.filter(pl.col("n_innovation") == total_max).select("year").item()
)
bioeconomy_max = pct_bioeconomy.select("n_bioeconomy").max().item()
bioeconomy_max_year = (
    pct_bioeconomy.filter(pl.col("n_bioeconomy") == bioeconomy_max)
    .select("year")
    .item()
)

total_pct_bioeconomy = sum_bioeconomy_innovations / n_swinno

```

```{python}
# | label: Number of Results
# | output: asis

count_bioeconomy_visions = bioeconomy_visions.group_by(
    ["year", "bioeconomy_vision"]
).agg(pl.col("sinno_id").count().alias("count"))

total_bioeconomy_visions = (
    count_bioeconomy_visions.group_by("bioeconomy_vision")
    .agg(pl.sum("count").alias("count"))
    .sort("count", descending=True)
)

n_bioeconomy_visions = bioeconomy_visions.select("sinno_id").n_unique()
bioeconomy_visions_min_year = bioeconomy_visions.select(pl.min("year")).item()
bioeconomy_visions_max_year = bioeconomy_visions.select(pl.max("year")).item()
n_most_bioeconomy_vision = total_bioeconomy_visions.row(0)[1]

def calculate_share(vision: str):
    return (
        total_bioeconomy_visions.filter(pl.col("bioeconomy_vision") == vision)
        .select("count")
        .item()
        / n_bioeconomy_visions
    )

pct_vision_neutral = calculate_share("Vision Neutral")
pct_bioecology = calculate_share("Bioecology Vision")
pct_bioresource = calculate_share("Bioresource Vision")
pct_biotechnology = calculate_share("Biotechnology Vision")

n_bioresource_1970_1990 = (
    bioeconomy_visions.filter(
        (pl.col("year").is_between(1970, 1990))
        & (pl.col("bioeconomy_vision") == "Bioresource Vision")
    )
    .select("sinno_id")
    .n_unique()
)

n_bioresource = (
    bioeconomy_visions.filter(pl.col("bioeconomy_vision") == "Bioresource Vision")
    .select("sinno_id")
    .n_unique()
)

n_biotechnology = (
    bioeconomy_visions.filter(pl.col("bioeconomy_vision") == "Biotechnology Vision")
    .select("sinno_id")
    .n_unique()
)

n_bioecology = (
    bioeconomy_visions.filter(pl.col("bioeconomy_vision") == "Bioecology Vision")
    .select("sinno_id")
    .n_unique()
)

n_vision_neutral = (
    bioeconomy_visions.filter(pl.col("bioeconomy_vision") == "Vision Neutral")
    .select("sinno_id")
    .n_unique()
)
```

```{python}
# | label: Count Recycling Innovation

recycle_innovation = pl.read_csv(RAW_DATA_DIR / "recycling_innovation.csv")

recycle_innovation_count = recycle_innovation["sinno_id"].n_unique()

```